Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-06 ms.tex
section 5

Figure~\ref{fig:vwSources} shows $\vwO(M_{\bullet})$ from each heating
source (stars, MSPs, SNe Ia, and black hole feedback) calculated for
the average star formation history of galaxies containing a given
black hole mass.  Again, Ia heating is suppressed in low mass galaxies
because the Ia radius exceeds the nominal stagnation radius (bottom
panel).  MSP heating is again found to be negligible for all black
hole masses.  The younger stellar populations characterizing galaxies
with $\Mbh\lsim 5\times 10^{8} \Msun$ are dominated by stellar wind
and Ia heating, with $v_{w}^{\star} \sim 1000$ km s$^{-1}$ and
$v_w^{\rm Ia} \sim 500$ km s$^{-1}$, respectively.  Only for the most
massive galaxies, with $\Mbh\gsim 5\times 10^{8} \Msun$, does the lack
of young stellar populations significantly reduce the role of stellar
wind
heating.  %Solid and dashed lines again show black hole masses for which the flow is thermally stable and thermally unstable, respectively, with $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}}$ as a function of $M_{\bullet}$ shown in the bottom panel.
Strikingly, {\it we find that galactic nuclei experiencing the same
  average star formation history as their host galaxies produce
  thermally stable flows across almost all black hole masses}.  One important exception occurs for the highest mass black holes, for which the total effective wind velocity is comparable to the stellar velocity dispersion ($\zeta \sim \zeta_c$) and thermal stability for the average star formation history is marginal, especially for core galaxies ($\S\ref{sec:mdot}$).

Realistic variations (``burstiness") in the star formation history
will often result in lower heating rates.  Figure \ref{fig:NickPlot2}
shows the heating rate as a function of average black hole mass for
non-fiducial cases in which the current ($z = 0$) star formation is
suppressed by a factor $\iota$ from its average $z = 0$ value, for a
characteristic timescale $\delta t_{\star}$ (for details, see Appendix
\ref{app:windheat}).  For $\delta t_{\star} \lesssim 10^{7}$ yr, we see
that $v_{\rm w}^{\star}$ is reduced by at most a factor of $\approx 2$
from its average value, even for huge drops in the star formation rate
($\iota \sim 10^{-3}$).  However, burstiness in the star formation
history over longer timescales can suppress heating more
significantly.  When $\delta t_{\star} \gtrsim 10^8~{\rm yr}$, $v_{\rm
  w}^\star$ becomes very sensitive to $\iota$: at $\iota=0.1$, $v_{\rm
  w}^\star \approx 500~{\rm km~s}^{-1}$ in most galaxies, but for
smaller $\iota$, $v_{\rm w}^\star \approx 200~{\rm km~s}^{-1}$ (an
exception to this is in galactic nuclei with $M_\bullet \gtrsim
5\times 10^8~M_\odot$, where heating is stabilized by SN Ia).  Our use
of average cosmic star formation histories is appropriate provided
local variations are either on short timescales ($\delta t_{\star}
\lesssim 10^7~{\rm yr}$), or limited in magnitude ($\iota \gtrsim
0.1$).  If both of these conditions are violated, then the effective
heating rates for all but the most massive galaxies (where Ia
explosions dominate) will fall by a factor $\approx 5$ from our
fiducial volumetric averages, taken from \citet{MosterNaab+:2013a}.

>>>>

Figure~\ref{fig:vwSources} shows $\vwO(M_{\bullet})$ from each heating
source (stars, MSPs, SNe Ia, and black hole feedback) calculated for
the average star formation history of galaxies containing a given
black hole mass.  The younger stellar populations characterizing galaxies
with $\Mbh\lsim 3\times 10^{8} \Msun$ are dominated by stellar wind, with $v_{w}^{\star} \gtrsim 700$ km s$^{-1}$.  Only for the most massive galaxies, with $\Mbh\gsim 3\times 10^{8} \Msun$, does the lack
of young stellar populations significantly reduce the role of stellar
wind heating.  But for these massive galaxies, Ia heating contributes a comparable level of heating $v_{w}^{\rm Ia} \gtrsim 600$ km s$^{-1}$.  However, for low mass galaxies Ia heating does not contribute to the total heating rate because the Ia radius exceeds the nominal stagnation radius (bottom panel).  Finall, MSP and Compton heating is found to be negligible for all black hole masses.  Strikingly, {\it we find that galactic nuclei experiencing the same
  average star formation history as their host galaxies produce thermally stable flows across almost all black hole masses}.  
We must caveat, however, that for high $M_{\bullet}$ the total effective wind velocity is only a factor of 2-3 times higher than the stellar velocity dispersion ($\zeta \sim \zeta_c$), such that thermal stability is marginal, especially for core galaxies ($\S\ref{sec:mdot}$).

Realistic variations (``burstiness") in the star formation history
will also often result in lower heating rates (Appendix
\ref{app:windheat}).  Figure \ref{fig:NickPlot2}
shows the heating rate as a function of average black hole mass for
non-fiducial cases in which the current ($z = 0$) star formation is
suppressed by a factor $\iota$ from its average $z = 0$ value, for a
characteristic timescale $\delta t_{\star}$.  For $\delta t_{\star} \lesssim 10^{7}$ yr, we see
that $v_{\rm w}^{\star}$ is reduced by at most a factor of $\approx 2$
from its average value, even for huge drops in the star formation rate
($\iota \sim 10^{-3}$).  However, burstiness in the star formation
history over longer timescales can suppress heating more
significantly.  When $\delta t_{\star} \gtrsim 10^8~{\rm yr}$, $v_{\rm
  w}^\star$ becomes very sensitive to $\iota$: at $\iota=0.1$, $v_{\rm
  w}^\star \approx 500~{\rm km~s}^{-1}$ in most galaxies, but for
smaller $\iota$, $v_{\rm w}^\star \approx 200~{\rm km~s}^{-1}$ (an
exception to this is in galactic nuclei with $M_\bullet \gtrsim
5\times 10^8~M_\odot$, where heating is stabilized by SN Ia).  Our use
of average cosmic star formation histories is appropriate provided
local variations are either on short timescales ($\delta t_{\star}
\lesssim 10^7~{\rm yr}$), or limited in magnitude ($\iota \gtrsim
0.1$).  If both of these conditions are violated, then the effective
heating rates for all but the most massive galaxies (where Ia
explosions dominate) will fall by a factor $\approx 5$ from our
fiducial volumetric averages, taken from \citet{MosterNaab+:2013a}.
